---
title: "Example of Cramer's V Calculation in R"
author: "Jasmine Dumas"
date: "February 23, 2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


## Usage and interpretation

In statistics, Cramer's V (sometimes referred to as Cramer's phi and denoted as $\phi$ c) is a measure of association between two nominal variables, giving a value between 0 and +1 (inclusive). It is based on Pearson's chi-squared statistic and was published by Harald Cramer in 1946. $\phi$ c is the intercorrelation of two discrete variables. Cramer's V varies from **0** (corresponding to no association between the variables) to **1** (complete association) and can reach 1 only when the two variables are equal to each other. $\phi^2$ c is the **mean square canonical correlation** between the variables. Source: <https://en.wikipedia.org/wiki/Cram%C3%A9r%27s_V>

_______

# airquality dataset

```{r}
library(lsr)

head(airquality)

airquality <- na.omit(airquality)

library(dplyr)

olddat <- airquality %>%  filter(Month < 7) # split the data in two groups by month 

newdat <- airquality %>% filter(Month >= 7)

old_cramers <- apply(olddat, 2, cramersV) # 2 indicates apply the function over the columns

new_cramers <- apply(newdat, 2, cramersV)

sort(old_cramers, decreasing = TRUE) # highest values first

sort(new_cramers, decreasing = TRUE)

old_new <- cbind(old_cramers, new_cramers) # bind the columns

```


## Static graphics with base
```{r}
# http://thecoatlessprofessor.com/programming/creating-stacked-barplot-and-grouped-barplot-in-r-using-base-graphics-no-ggplot2/
par(mar=c(5.1, 4.1, 4.1, 7.1), xpd=TRUE)
prop = prop.table(old_new,margin=2)

barplot(old_new, col=heat.colors(length(rownames(old_new))), main = "Stacked count")
legend("topright",inset=c(-.25,0), fill=heat.colors(length(rownames(old_new))), legend=rownames(old_new))

# unstacked barplot
barplot(prop, col=heat.colors(length(rownames(prop))), width=2, beside=TRUE, main = "Unstacked count")
legend("topright",inset=c(-.25,0), fill=heat.colors(length(rownames(prop))), legend=rownames(old_new))

# stacked percentage based
barplot(prop, col=heat.colors(length(rownames(prop))), width=2, main = "Stacked %")
legend("topright",inset=c(-0.25,0), fill=heat.colors(length(rownames(prop))), legend=rownames(old_new))

```

## Dynamic grpahic with plotly
```{r}
library(plotly)
library(reshape2)

# suggested layout from Paul to clearly see differences
p = melt(data = prop, 
                    measure.vars = c("Va1", "Var2", "value"))
as.list(p)

plot_p = plot_ly(p, 
                x = Var1, y = value, 
                color = factor(Var2), 
                type ="bar")
plot_p


```

