---
title: "Automated Variable Selection Algorithms"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
layout: page
---


```{r setup, include=FALSE}
# to be rendered as html for papers.md section

knitr::opts_chunk$set(fig.path='{{ site.url }}/post_data/auto-vars-algos-')
```


# Get the data
```{r}
load("~/Desktop/depaul/CSC423/rdata/R/Exercises&Examples/EXEXSAL2.Rdata")
```

# Backward Elimination Selection

* Starts will all variables.
* Drops one variable at a time until dropping another variable no longer improves the model.
* Once a variable is dropped it cannot re-enter the model.
* With AIC criteria (which R uses here) lower is better. Note: lowest may include most negative.
* AIC = Akaike's Information Criteria = n * ln(SSE/n) + 2 * p

```{r}
library(MASS)
full.model <- lm(Y~X1+X2+X3+X4+X5+X6+X7+X8+X9+X10, data=EXEXSAL2)
model = step(full.model, direction="backward")
summary(model)
```

# Forward Selection

* Adds one variable at a time until adding a new variable no longer improves the model.
* Once a variable is added, it never leaves the model.
* With AIC criteria (which R uses here) lower is better.

```{r}
min.model = lm(Y ~ 1, data=EXEXSAL2)          # aka intercept only model
biggest = formula(lm(Y ~ ., data=EXEXSAL2))   # note dot: all variables, including ID
model = step(min.model, direction='forward', scope=biggest)
summary(model)
```

# Stepwise Regression Selection

* Adds one variable at a time until adding a new variable no longer improves the model.
* A variable may leave and re-enter the model many times.
* With AIC criteria (which R uses) lower is better.
* Note I am reusing full.model here.

```{r}
model = step(full.model, direction="both")
summary(model)
```

# Best-N Subset aka All Possible Regressions Selection

* Use leaps library.
* nbest shows n best models for each k predictors, for at most n*(k-1)+1 models
* For Mallow's Cp, want Cp “small and near” p (recall p = k + 1) (See slides, quote from SAS Support about Cp)
* Can use method=“Cp” then model$Cp, or method=“adjr2” then model$adjr2

## Best Subset using Cp

```{r}
library(leaps)
yvar = c("Y")
xvars = c("X1", "X2", "X3", "X4", "X5", "X6", "X7", "X8", "X9", "X10")
model=leaps( x=EXEXSAL2[,xvars], y=EXEXSAL2[,yvar], names=xvars, nbest=3, method="Cp")
model$which
model$Cp
```

## Best Subset using adjr2

```{r}
yvar = c("Y")
xvars = c("X1", "X2", "X3", "X4", "X5", "X6", "X7", "X8", "X9", "X10")
model=leaps( x=EXEXSAL2[,xvars], y=EXEXSAL2[,yvar], names=xvars, nbest=3, method="adjr2")
model$which
model$adjr2
```


# Additional Recource

<http://billqualls.com/csc423/csc423_ex_06_01_r.html>



