---
layout: post
title: "Extracting data from Twitter for @hrbrmstr's #nom foodie images"
output: 
  html_document: 
    self_contained: no
tags: [rstats, r, r-bloggers, tutorial, food]
bigimg: /post_data/data/Sage rosemary &amp; espresso infused salt rubbed roast lamb. Goose fat roasted potatoes _almost _ done .jpg
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE

)
```

[Bob Rudis](https://rud.is/) (@hrbrmstr) is a famed expert, author and developer in Data Security and the Chief Security Data Scientist at Rapid7. Bob also creates the most deliciously vivid images of his meals documented by the #nom hashtag. I'm going to use a similar method used in my previous projects ([Hipster Veggies](https://github.com/jasdumas/hipster-veggies) & [Machine Learning Flashcards](https://jasdumas.github.io/2017-05-02-twitter-ml-flashcards/)) to wrangle all those images into a nice collection - mostly for me to look at for inspiration in recipe planning.

<blockquote class="twitter-tweet" data-lang="en"><p lang="en" dir="ltr">Yum! Have you ever thought about collecting all these recipes &amp; images into a cookbook?!</p>&mdash; Jasmine Dumas (@jasdumas) <a href="https://twitter.com/jasdumas/status/952971103990898689?ref_src=twsrc%5Etfw">January 15, 2018</a></blockquote>
<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

Source Repository: [jasdumas/bobs-noms](https://github.com/jasdumas/bobs-noms)

## Analysis

```{r message=FALSE, warning=FALSE, libs, message=FALSE}
library(rtweet) # devtools::install_github("mkearney/rtweet")
library(tidyverse)
library(dplyr)
library(stringr)
library(magick)
library(knitr)
library(kableExtra)
```


```{r, get_tweets}
# get all of bob's recent tweets
bobs_tweets <- get_timeline(user = "hrbrmstr", n = 3200)

#filter noms with images only
bobs_noms <- 
  bobs_tweets %>% dplyr::filter(str_detect(hashtags, "nom"), !is.na(media_url))
```


```{r, clean_text}
bobs_noms$clean_text <- bobs_noms$text
bobs_noms$clean_text <- str_replace(bobs_noms$clean_text,"#[a-zA-Z0-9]{1,}", "") # remove the hashtag
bobs_noms$clean_text <- str_replace(bobs_noms$clean_text, " ?(f|ht)(tp)(s?)(://)(.*)[.|/](.*)", "") # remove the url link
bobs_noms$clean_text <- str_replace(bobs_noms$clean_text, "[[:punct:]]", "") # remove punctuation
```


```{r, view_imgs}
# let's look at these images in a smaller data set
bobs_noms_small <- bobs_noms %>% select(created_at, clean_text, media_url)

bobs_noms_small$img_md <- paste0("![", bobs_noms_small$clean_text, "](", bobs_noms_small$media_url, ")")
```

```{r, imgs, asis=T}
bobs_noms_small$img_md %>% 
kable( format = "markdown") %>%
  kable_styling(bootstrap_options = "striped", 
                full_width = F) 
```

```{r, func}
# create a function to save these images!
save_image <- function(df){
  for (i in c(1:nrow(df))){
    image <- try(image_read(df$media_url[[i]]), silent = F)
  if(class(image)[1] != "try-error"){
    image %>%
      image_scale("1200x700") %>%
      image_write(paste0("../post_data/data/", bobs_noms$clean_text[i],".jpg"))
  }
 
  }
   cat("saved images...\n")
}

save_image(bobs_noms)
```